---
title: "Avian Response to Recovering Landscapes"
author: "James Waterford"
date: "2023-04-05"
output: html_document
---

Avian functional responses to landscape recovery
---
Citation: Ikin, Karen et al. (2019), Data from: Avian functional responses to landscape recovery, Dryad, Dataset, https://doi.org/10.5061/dryad.d6k0804

This data set has been retrieved from the DRYAD public domain, and released by Charles Darwin University, McGill University, and James Cook University.

The aim of this database was to measure the effect of restoring vegetation across **17** sites in Australia.
This data set also includes the ecological traits of these birds, and the attributes of the restored sites.



Background
---



After searching DRYAD for an explanation of the dataset, I found a supplementary material on the published article in the [Journal of Royal Society Publishing](https://rs.figshare.com/articles/journal_contribution/ALL_FILES_IN_ONE_PDF_-_Table_S1_Bird_species_and_assigned_guilds_Table_S2_Model_selection_functional_and_taxonomic_richness_landscape_recovery_Table_S3_Model_selection_functional_richness_taxonomic_richness_Table_S4_Model_summaries_functional_and_taxonomi/7937501?backTo=/collections/Supplementary_material_from_Avian_functional_responses_to_landscape_recovery_/4455659).


```{r setup, echo=FALSE, message=FALSE}
library(dplyr)
library(stringr)
library(ggplot2)
library(ggmap)
```


These first two data files were imported directly from DRYAD. 


```{r import}

BirdData_full <- read.delim("../data-raw/Ikin_2019_BirdData.txt")
SiteData_full <- read.delim("../data-raw/Ikin_2019_SiteData.txt")

head(colnames(BirdData_full))
```


The files from DRYAD have abbreviated the species names, so I combined them with the supplemental material to recreate the table.


```{r}
species <- read.csv("../data-raw/SPECIES-GENUS.csv")
new_names <- c("id")
append(x = new_names,species$Species,1) -> new_names
colnames(BirdData_full) <- c(new_names)

BirdSite_Data <- full_join(BirdData_full, SiteData_full, by ="id")
apply(X = BirdSite_Data[2:122], MARGIN = 2, sum) -> small_table
```

The *column names* for the bird data reference different species that are local to Australia.

The **IDs** reference multiple measurements at multiple sites.

The `SiteData.txt` has all the information on the various sites, including their geographic location.

```{r}
# BirdSite_Data <- full_join(BirdData_full, SiteData_full, by ="id")
head(BirdSite_Data[122:130])
```
Across New South Wales
---
Below is a heatmap of the sites that were measured.
```{r ggmap, echo=FALSE, warning=FALSE,message=FALSE}
library(ggmap)
library(RColorBrewer)
map_bounds <- c(left = 135, bottom = -39, right = 155, top = -18)
coords.map <- get_stamenmap(map_bounds, zoom = 5, maptype = "toner-lite")
coords.map <- ggmap(coords.map, extent="device", legend="none")
coords.map <- coords.map + stat_density2d(data=SiteData_full,  aes(x=long, y=lat, fill=..level.., alpha=..level..), geom="polygon")
coords.map <- coords.map +   scale_fill_gradientn(colours=rev(brewer.pal(7, "Spectral")))
coords.map
```

The sites are in the southern region of New South Wales, a state that is currently deforesting portions of the **Nagmadi National Forest** in Canberra.

While this dataset only is considering the family of birds in NSW, it is relevant to see how deforesting can affect native species populations.
```{r ggmap2, echo=FALSE, warning=FALSE,message=FALSE}
map_bounds <- c(left = 145, bottom = -36.5, right = 150, top = -34.5)
coords.map <- get_stamenmap(map_bounds, zoom = 8, maptype = "toner-lite")
coords.map <- ggmap(coords.map, extent="device", legend="none")
coords.map <- coords.map + stat_density2d(data=SiteData_full,  aes(x=long, y=lat, fill=..level.., alpha=..level..), geom="polygon")
coords.map <- coords.map +   scale_fill_gradientn(colours=rev(brewer.pal(7, "Spectral")))
coords.map
```

The `SiteData.txt` file also includes background on the physical state of the farm (Planting / Regrowing / No Growth), as well as the tree coverage.

Both of these are important vectors to consider the effect on the species.

<!-- Explanation of how we got to big table, and how this apply function works. -->


Initial Explanation of Variables
---

```{r ggplot2}
BirdSite_Data %>% 
  filter(ln_cover != 0) -> BirdSite_Data


ggplot(data=BirdSite_Data, mapping = aes(x = farm, y = site.rich, color = growth)) +
  geom_point(size = 2.25) +
  geom_smooth(method = lm) +
  xlab(label = "Active Sites in New South Wales") +
  ylab(label = "Total count per Visit") 


ggplot(data=BirdSite_Data, mapping = aes(x = ln_cover, y = site.rich, color = growth)) +
  geom_point() +
  geom_smooth(method = loess) +
  xlab(label = "Ln of tree coverage per site") +
  ylab(label = "Total count per visit")
```

As we can see from the dot plots, there is a greater concentration of species on sites where significant regrowth has occurred. 

<!-- Come back and edit your first Dotplot -->

Taxonomic Analysis & OpenTree Taxonomy
-------------------

Looking at the interconnectivity of the species present across these NSW will help us better understand the ecological variety.

<!-- Describe why we are using Opentree -->

**OpenTree** is a taxonomic package that will allow us to to sort species by scientific name into a combined taxon tree.

The author had included a supplementary data table with their original [article](https://rs.figshare.com/articles/journal_contribution/ALL_FILES_IN_ONE_PDF_-_Table_S1_Bird_species_and_assigned_guilds_Table_S2_Model_selection_functional_and_taxonomic_richness_landscape_recovery_Table_S3_Model_selection_functional_richness_taxonomic_richness_Table_S4_Model_summaries_functional_and_taxonomi/7937501?backTo=/collections/Supplementary_material_from_Avian_functional_responses_to_landscape_recovery_/4455659).

With this dataset,

```{r OpenTree_load, echo=FALSE, warning=FALSE}
library(ggtree)
library(ape)
library(rotl)
```

I am reducing the number of species in this table to just the genuses with counts greater than 150.

```{r, warning=FALSE}
small_table <- data.frame((small_table))

colnames(small_table)[1] <- "Count"
small_table$Species <- c(species$Species[1:121])
small_table %>% 
  filter(Count >= 150) -> smaller_table

left_join(smaller_table,species, by = "Species") -> bird_list

bird_taxa <- tnrs_match_names(bird_list$Scientific.name)

bird_tr <- tol_induced_subtree(ott_ids = bird_taxa$ott_id)
plot(bird_tr)
```



