---
title: "Visualization in R"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualiztion in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(readr)
library(ggplot2)
```

What is the purpose of this page?
----------------------



This is a short tutorial on the basics of **extracting**, **evaluating**, and **plotting** in RStudio.



To do this best, we will be working with a practice dataset from The Ecological Society of America, from a paper entitled **Plant and small-mammal responses to large-herbivore exclusion in an African savanna: five years of the UHURU experiment**.

UHURU studies the effect of herbivores on various plants in Kenya.

This data set in particular focuses on the effects of different *treatments* on Acacia trees.
These *treatments* are different areas of Kenya, where predation levels (and species) are diversified.


We will first call the file, based on the current directory. Then we will use `read.table()` to access the file. 

``` {r Pull & Create tsv, eval  = TRUE}
getwd()
## Provide file using a relative path ##
acacia_csv <- ("../data-raw/ACACIA_DREPANOLOBIUM_SURVEY.txt")
## add `na.strings = "dead"
acacia <- read.table(acacia_csv, header = TRUE, sep = "\t", fill = TRUE)
```

Modifying a data frame
-------------------
We've read the file into  a variable in our environment. 

What can we quickly find out from this data set?

There are a handful of commands that can help us out. Some personal favorites are:


 `head` - read the first 10 lines from the first 10 columns
 `summary` - does simple mathematical calculations on your data set
 `str` - displays 10 values from each column, with their data type
 `View` - creates a table in another tab
 `class` - disaplys the data type
 
```{r read through, eval = TRUE}
#head(acacia)
str(acacia)
#summary(acacia)
```
After using the `str()` function, we see some small oddities in our data set. 

For example, *Why is height a character vector?* This can be an issue for some of the mathematical calculations. 

Let's try and find where the issue lies.

For Data Frames, we sort by Column, Then Row By using `df[Column,Row]`, where `df` refers to the name of your data frame.

If you run `df[,X]`, you will get the full list of values from a Column.

We can also use `$` as to sort through columns.

Since we know that some aspect of `HEIGHT` is leading to this column being a `chr` type, let's look at the full column to see where the issue lies.
```{r}

acacia$HEIGHT
acacia[21,8]

```
It appears that the authors of this dataset referred to deceased plants as 'dead'. However, if we want `HEIGHT` as a `num` type, we will need to replace those values.

One function that does this, in the simplest method, is `as.numeric()`.

We will create a new vector that can replace the column in our data frame.
``` {r replace}
col_height <- as.numeric(acacia$HEIGHT)
acacia$HEIGHT <- col_height
```

By running this command, we have *coerced* the character strings into NA values. This way, we can run mathematical statistics on the data set.

If we caught this earlier, maybe before we imported the data set, we could've forced NAs during import. Let's try this now with a feature from  `read.table()`, referred to as `na.strings = `
``` {r}
acacia <- read.table(acacia_csv, header = TRUE, sep = "\t", na.strings = "dead")
```


Can we make a graph from this data? 
-----------------------------------

Now that we have a clean dataset, let's try plotting it with `plot()`.
```{r graph step 1}
plot(x=acacia$HEIGHT,y=acacia$AXIS1)
```

The `plot` function is useful, but not powerful. Its hard to add titles, and even more difficult to change plot types. How could you ever do a heatmap via plot? You can't. So instead we use `ggplot`.

How do we use ggplot?
----------------------

`ggplot2` is the standard R package used for graphing, and it is highly modifiable.

The basic setup for a ggplot is as follows:

`ggplot(data = df, mapping = aes(x = [independent_variable]), y = [dependent_variable] ) + `
  `geom_[graph_type] + `
  `...addtl args `

We will use `geom_point()` to create a scatter plot, measuring `HEIGHT` against `AXIS1` (width)
```{r graph step 2}
library(ggplot2)
ggplot(data=acacia, mapping = aes(x=HEIGHT, y= AXIS1, color = TREATMENT)) +
  geom_point() +
  labs(x="Tree Height", y = "Tree Axis 1")
```

What if we want to reshape our axes? 
We have functions to `scale()` the data set.
```{r graph step 3}
ggplot(data = acacia, mapping = aes(x = HEIGHT, y = FRUITS, color = ANT)) +
  geom_point(size = 3, alpha= 0.5) +
## Use facet_wrap 
  facet_wrap(~ANT, scales = "free")
  
## Add in geom_smooth() for data insigh t##
  
ggplot(data = acacia, mapping = aes(x = CIRC, y = HEIGHT, color = TREATMENT)) +
  ## alpha is a modifier of point transparaceny
  geom_point(size = 3, alpha = 0.667) + 
  geom_smooth() #method = " +
  ggsave(filename = "jw_acacia_treatment.jpg")

```
You must always call `ggplot` to access any of its graphical interface. The differences lie in the interlaced functions.

```{r graphs3_4}
ggplot(data=acacia, mapping = aes(x=TREATMENT)) +
  geom_bar()
```
```{r}
ggplot(acacia, aes(x=CIRC, color= TREATMENT)) +
  ## `bins = ` defines how many boxes are displayed 
  ## `Fill = ` is for color 
  geom_histogram(bins = 20, fill = "slategray3")
  #ggsave("../197-figures/acacia_Circ_by_treatment.jpg")
```

Add an example of Geom_line here :)
